### 2013-08-11 apie-2013/paper.tex
Added skeletons for api and iced papers to be written.

\documentclass[conference]{IEEEtran}

% Use of outside images
\usepackage{graphicx} 
% Use text inside euqations
\usepackage{amsmath}

\usepackage{balance}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}

% Correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

% Begin the paper here
\begin{document}

% Paper title
% Can use linebreaks \\ within to get better formatting as desired
\title{APIE Evolution: An Exploration into Software API Trends and Processes}

% Authors names
\author{\IEEEauthorblockN{Jordan Ell}
\IEEEauthorblockA{University of Victoria,
Victoria, British Columbia, Canada \\ jell@uvic.ca}
}

\author{\IEEEauthorblockN{Braden Simpson}
\IEEEauthorblockA{University of Victoria,
Victoria, British Columbia, Canada \\ braden@uvic.ca}
}

\author{\IEEEauthorblockN{Daniela Damian}
\IEEEauthorblockA{University of Victoria,
Victoria, British Columbia, Canada \\ danielad@csc.uvic.ca}
}

% Make the title area
\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Related Work}

\section{Methodology}

\subsection{Results}

\section{Discussion}

\section{Future Work}

\section{Conclusions}

\bibliographystyle{IEEEtran}
\balance
\bibliography{paper}

% End of the paper
\end{document}