### 2013-12-03 msc-thesis/chapters/1/chapter-intro.tex
MSC Added in 2 motivational studies.

\startfirstchapter{Introduction}
\label{chapter:introduction}

\input chapters/1/sec-intro
\input chapters/1/sec-review