### 2013-12-03 msc-thesis/chapters/2/chapter-problem.tex

\startchapter{The Problem}
\label{chapter:problem}

\newlength{\savedunitlength}
\setlength{\unitlength}{2em}

Introduction paragraph about the problem and the various motivations.

\input chapters/2/sec-developer-pairs
\input chapters/2/sec-impact

\section{Problem Description}

Use the two case studies as motivating examples to show the full problem.

\setlength{\unitlength}{\savedunitlength}