### 2013-12-03 msc-thesis/chapters/3/chapter-new.tex

\startchapter{The New Approach and Solution}
\label{chapter:newsol}

\input chapters/3/sec-latexhelp