### 2014-01-05 msc-thesis/chapters/4/chapter-discussion.tex

\startchapter{Discussion}
\label{chapter:discussion}