### 2014-01-05 msc-thesis/frontmatter/abs.tex

\newpage
\TOCadd{Abstract}

\noindent \textbf{Supervisory Committee}
\tpbreak
\panel

\begin{center}
\textbf{ABSTRACT}
\end{center}

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused 
on either \textit{direct} or 
\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
from a result of source code changes. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied for indirect conflicts have had common issues of
information overload, false positives, scalability, information distribution and many others. 
To better understand these issues, this dissertation will focus on exploring the world of indirect conflicts
through 4 studies. The first two studies presented will focus on motivational circumstances which occur
during the software development life cycle and cause indirect conflicts. Developers interactions are
studied in order to create a tool which can aid in the workflows around indirect conflicts. The second
two studies present a deeper investigation into why most indirect conflict tools fail to attrack developer
inerest through exploring the root causes of indirect conflicts and how tools should be properly built
to support developer workflows.