### 2014-03-16 iced-2013/paper.tex
ICED final commit.

\documentclass[conference]{IEEEtran}

% Use of outside images
\usepackage{graphicx}
% Use text inside euqations
\usepackage{amsmath}

%\usepackage{balance}
\usepackage{float}

\usepackage{hyperref}
\usepackage{nameref}

\floatstyle{plaintop}
\restylefloat{table}

% Correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\makeatletter
\def\namedlabel#1#2{\begingroup
   \def\@currentlabel{#2}%
   \label{#1}\endgroup
}
\makeatother

\newcommand{\totalCategories}{10}

% Begin the paper here
\begin{document}

% Paper title
% Can use linebreaks \\ within to get better formatting as desired
\title{State of The Practice in Indirect Conflicts: Causes, Mitigation Strategies, and Future Directions}

\author{\IEEEauthorblockN{Jordan Ell and Daniela Damian}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ jell@uvic.ca, danielad@cs.uvic.ca}
}

% Make the title area
\maketitle

\begin{abstract}

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused
on either \textit{direct} or
\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
from a result of source code changes. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied for indirect conflicts have had common issues of
information overload, false positives, and scalability. To better understand these issues,
we performed a mixed method study involving interviews with 19 developers from 12 commercial and open
source projects, and a questionnaire with 78 respondents different from our interviewees. Using a grounded theory approach to data analysis,
we obtained insights about events and conditions in which indirect conflicts are likely to occur in
software development, that developers prefer to use detection and resolution processes or tools
over those of prevention, that they do not want awareness mechanisms which provide non actionable results,
and that there exists a gap in software evolution analytical tools.

\end{abstract}

\section{Introduction}
\label{sec:intro}

As Software Configuration Management (SCM) has grown, the maturity and norm of parallel
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
empirical evidence shows that despite of recent research has proposed mitigation strategies, these losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL, Sarma:2009:TIV}.

Two types of conflicts have attracted the attention of researchers, \textit{direct} and
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and
positive developer feedback. Indirect conflicts (ICs) are caused by source code changes that negatively impact another
location in the software system such as when libraries are upgraded or when a method is changed
and invoking methods are influenced negatively. Indirect conflict tools however, have
not shared the same success as direct conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}.

While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, which causes tools to eventually be
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of code dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success; however, the
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers or unit tests, to these dependencies still remains an issue.
Some tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA} have also leveraged the notion of socio-technical congruence ~\cite{Cataldo:2006:ICR} to help developers pursue dependencies through social means.
However, the same aforementioned problems still arise. Evaluations of these tools or approaches have largely been small in scale, and involved a limited number of industrial participants or student evaluators. As a result, the problems of scalability, information overload and false positives in tackling indirect conflicts remains an outstanding research problem.

In this paper we report from a field study of software practitioners that sought to characterize the state-of-the-practice in indirect conflicts, particularly their causes and mitigation strategies. Our intention has been to uncover the root causes of information overload, false positives, and
scalability issues, in order to better understand why indirect conflict tools and methods fail to fully achieve
their intended success. Three research questions that guided our study:

\begin{description}
	\item[RQ1\namedlabel{itm:rq1}{RQ1}] \textit{What are the types, factors, and frequencies of indirect conflicts?}
	\item[RQ2\namedlabel{itm:rq2}{RQ2}] \textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
	\item[RQ3\namedlabel{itm:rq3}{RQ3}] \textit{What do developers want from future indirect conflict tools?}
\end{description}

To answer these questions, we interviewed 19 developers from across 12 commercial and open source projects, sought confirmation of the themes we identified in our interviews through a questionnaire with 78
developers, as well as 5 follow-up interviews. Our findings indicate that:
\begin{itemize}
	\item Indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of such changes
	\item Developers tend to prefer to use detection and resolution processes or tools over those of prevention
	\item Developers do not want awareness mechanisms which provide non actionable results
	\item There exists a gap in software evolution analytical tools from the reliance on static analysis resulting in missed
				context of indirect conflicts
\end{itemize}

\section{Related Work}
\label{sec:related}

Many indirect conflict tools have been developed, tested, and published. Sarma et al.~\cite{Sarma:2007:TSA} created Palantir,
which can both detect potential indirect conflicts, at the class signature level, and alert developers to these conflicts.
Palantir represented one of the first serious attempts at aiding developers towards indirect conflicts. Holmes et
al.~\cite{Holmes:2010:CAR} take it one step further with their tool YooHoo, by detecting fine grained source code changes,
such as method return type changes, and create a taxonomy for different types of changes and their proneness to cause
indirect conflicts. This tool and its taxonomy had a severely reduced false positive rate, however, the true positives
detected may already be detectable by current tools such as compilers and unit testing. The tool Ariadne~\cite{Trainer:2005:BGT}
creates an environment where developers can see how source code changes will affect other areas of a project at the
method level, using method call graphs, showing where indirect conflicts may occur. This type of exploratory design has
been used often in the visualization of indirect conflict tools, allowing developers a type of search area for their development
needs. Another indirect conflict tool, CASI~\cite{Servant:2010:CPI}, utilizes dependency slicing~\cite{Bajracharya:2009:SIS}
instead of method call graphs to provide an environment to see what areas of a project are being affected by a source code change.
The evaluations of these tools indicate that they have the same common difficulties of scalability, false positives, and information overload. Our own tool
Impact!~\cite{Ell:2013} also suffered these same fates.

Since Cataldo et al.~\cite{Cataldo:2006:ICR} have shown that socio-technical congruence can be leveraged to improve task completion
times, many indirect conflict tools support the idea of a socio-technical congruence~\cite{Kwan:2011:ESC} in order to help
developers solve their indirect conflict issues through social means~\cite{Begel:2010:CDE}~\cite{Borici:2012:CHA}.
Begel et al.~\cite{Begel:2010:CDE} created Codebook, a type of social developer network related directly to source code
artifacts, which can be used to identify developers and expert knowledge of the code base. Borici et al.~\cite{Borici:2012:CHA}
created ProxiScientia which used technical dependencies between developers to create a network of coordination needs.
However, again the problems of scalability and information overload become a factor.

For developer interest in regards to software modifications, Kim~\cite{Kim:2011:ESA} found that developers wanted to know whose
recent code changes semantically interfere with their own code changes, and whether their code is impacted by a code change.
Kim found that developers are concerned with interfaces of objects and when those interfaces change. Kim also identified the same
issues towards information overload through false positives with developers noting ``I get a big laundry list... I see the email and I delete it''.
Kim's field study does however fall short in actually creating a resolution for indirect conflicts, or finding new concerns of
developers which are not already detected by compilation or other static analysis tools.
In studying awareness of change impact, deSouza et. al.~\cite{deSouza:2008:ESS} found that developers use their personal knowledge of the
code base to determine the impact of their code changes on fellow developers, teams, and projects. However, their work does not
study which types of changes (types, frequencies, compounding factors) developers encounter in their work, nor provides concrete information on formal mitigation strategies or resolutions of indirect conflicts.

The insights of our study intends to fill the gap. In our field study we seek to understand not only  why information overload, false positives, and scalability are such difficult problems to tackle
in indirect conflict tools, but also how developers currently deal with indirect conflicts in practice through
their mitigation strategies, what are their major concerns as well as suggestions for future research and tools to tackle these concerns.

\section{Methodology}
\label{sec:meth}

We performed a mixed method study in three parts. First, a round of semi-structured interviews were conducted which
addressed our 3 main research questions. Second, a questionnaire was conducted
which was used to confirm and test what was theorized from the interviews on a larger sample size as well as obtain
larger developer opinion of the subject. Third, 5 confirmatory interviews were conducted by re-interviewing original
interview participants to once again confirm our insights.
We used grounded theory techniques to analyze the information provided from all three data gathering stages.

\subsection{Interview Participants}

Our interview participants came from a wide range of both open and closed source software development companies
and projects, using both  agile and waterfall based methodologies, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
The participants had software development experience in average of 8 years, as well as some project management experience.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| l | p{1cm} | p{2cm} | p{2cm} | p{2cm} | p{2cm} |}
\hline
Company & \# of Participants & Software Development Experience (years) & Development Process & Software Access & Current Language Focuses \\
\hline
\hline
Amazon & 2 & 5, 7 & Agile & Closed source & C++ \\ \hline
Exporq Oy & 1 & 8 & Agile & Closed source & Ruby, JavaScript \\ \hline
Fireworks Design & 1 & 6 & Agile & Closed source & JavaScript \\ \hline
Frost Tree Games & 1 & 4 & Agile & Closed source & C\# \\ \hline
GNOME & 1 & 13 & Agile & Open source & C \\ \hline
James Evans and Associates & 5 & 3, 3, 3, 4, 13 & Waterfall & Closed source & Oracle Forms \\ \hline
Kano Apps & 1 & 10 & Agile & Closed source & JavaScript, PHP \\ \hline
IBM & 2 & 5, 18 & Agile & Open and closed source & Java, JavaScript \\ \hline
Microsoft & 2 & 6, 10 & Agile & Closed source & C\# \\ \hline
Mozilla & 1 & 25 & Agile & Open source & C++, JavaScript \\ \hline
Ruboss & 1 & 5 & Agile & Closed source & JavaScript \\ \hline
Subnet Solutions & 1 & 5 & Agile & Closed source & C++ \\ \hline

\end{tabular}
\end{center}
\caption{Demographic information of interview participants.\label{tab:demo}}
\end{table*}

\subsection{Interview Procedure}

Participants were invited based on their software engineering and project management experience. We recruited participants through an email message to a number of software organizations in our working business network, and followed up in a single reminder email one week
after the initial invitation if no response was made. We directly emailed 22 participants and  conducted
19 interviews. Interviews were conducted in person when possible, or over video chat when not possible, and recorded for audio content only.

%We were pleased with the response rate to our initial invitation emails as
%our breadth of participants was rather large, spanning multiple open and closed source, Agile and Waterfall based projects,
%and our interview answer saturation~\cite{Guest:2006:SAP} was quite high.

Interview participants first answered a number of demographic questions and were then asked to describe various software development experiences and practices.
Specifically, ten semi-structured questions related to our research questions where ask and guided our interview:

\begin{itemize}
\item What tools are used for dependency tracking?
\item What processes are used for preventing ICs?
\item What artifact dependency levels are analyzed and where can conflicts arise?
\item How are software dependencies found?
\item Give examples of ICs from real world experiences.
\item How are ICs detected or found?
\item How are IC issues solved or dealt with?
\item Opinion of preemptive measures to prevent ICs.
\item What types of changes are worth a preemptive action?
\item Who is responsible for fixing or preventing ICs?
\end{itemize}

Not all participants had strong opinions or any experience on every category listed above. For these participants, answers
to the specific categories were not required. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being part of any one software project for extended periods of time.

\subsection{Questionnaire Participants}

Our questionnaire respondents were different from our interviewees. We recruited the respondents to our questionnaire from a similar
breadth of open and closed source software development
companies and projects as the interviews participants with two main exceptions. The GitHub users were recruited from the Google BigQuery database; the Apache respondents were recruited from the list of tops contributors to the Apache Software Foundation.
The software organizations that remained the same between interview and questionnaire were: Mozilla, The GNOME Project, Microsoft Corporation,
Subnet Solutions, and Amazon. The GitHub
users were selected based on large amounts of development activity on GitHub and the Apache developers
were selected based on their software development contributions on specific projects known to be used heavily
utilized by other organizations and projects.

\subsection{Questionnaire Procedure}

Questionnaire participants were invited to participate in the questionnaire by email. We directly emailed 1300 participants and obtained 78 responses
giving a response rate of 6\%. We attribute the low response rate with: the questionnaires
were conducted during the months of July and August while many participants may be away from their regular positions.

The questionnaire we designed ~\footnote{http://thesegalgroup.org/projects/indirect-conflicts/}
was based on the insights we obtained from the round of interviews, and was intended to confirm these insights but also to broaden them to a larger sample size of developers who may have similar or different opinions from those already acquired from the interviews. The question topics asked in the questionnaire were:

\begin{itemize}
\item What frequency do ICs occur at around different project milestones?
\item How does team size affect IC frequency?
\item What software change types do developers care about?
\item What processes are used for preventing ICs?
\item What tools are used for detecting ICs?
\item What tools are used for debugging ICs?
\end{itemize}

\subsection{Data Analysis}
To analyze both the interview and questionnaire data we used grounded theory techniques as described by Corbin and Strauss~\cite{Corbin:1998:SP}.
Grounded theory is a qualitative research methodology that utilizes \textit{theoretical sampling} and
\textit{open coding} to create a theory ``grounded'' in the empirical data. In this study we used grounded
theory techniques for analysis and not to develop a theory per say. For an exploratory study such as ours, grounded theory is well suited because it involves starting from very broad and abstract type questions, and
making refinements along the way as the study progresses and hypotheses begin to take shape. Grounded theory involves
realigning the sampling criteria throughout the course of the study to ensure that participants are able to answer new
questions that have been formulated in regards to forming hypotheses. In our study being presented, data collected from
both interviews and questionnaires (when open ended questions were involved) was analyzed using open and axial coding (relating
code and concepts to each other).
In Section~\ref{sec:eval}, we give a brief evaluation of our studying using
3 criteria that are commonly used in evaluating grounded theory studies.

\subsection{Validation}

Following our data collection and analysis, we re-interviewed 5 of our initial interview participants
in order to validate our findings. We confirmed our findings as to whether or not they resonate with
industry participants' opinions and experiences regarding indirect conflicts and as to their industrial
applicability. Due to limited time constraints of the interviewed participants, we could only re-interview
five participants. Those that were re-interviewed came from the range of 5-10 years of software development
experience. Re-interviewed participants were given our 3 research questions along with results and main
discussion points, and asked open ended questions regarding their opinions and experiences to validate our
findings. We also evaluated our grounded theory approach as per Corbin and Strauss'~\cite{Corbin:1998:SP}
list of criteria to evaluate quality and credibility, described in Section~\ref{sec:eval}.

\section{Results}
\label{sec:results}

In this section we present our results of both the interviews and questionnaires conducted in regards to our 3 research questions
outlined in Section~\ref{sec:intro}. For each research question we summarize the quantitative and qualitative
evidence we obtained. In each subsection, quantitative data
given refers to interviews conducted unless explicitly said otherwise.

\phantomsection
\addcontentsline{toc}{subsection}{RQ1}
\subsection*{\normalfont{RQ1.} \it{What are the types, factors, and frequencies of indirect conflicts?}}

\bf{The most common occurrence of indirect conflict is when a software object's contract changes. The frequency of
indirect conflicts, while usually high, decreases as a stable point is reached in the development cycle. The frequency
of indirect conflicts is compounded by the number of developers working on a project.}

\normalfont{}

We found that a large contributing factor to the cause
of indirect conflicts comes from the changing of a software object's contract (reported by 12 developers directly). Object contracts, or ``Design by Contract''\texttrademark, a concept made well-known through the work on Eiffel Software's~\footnote{http://www.eiffel.com/} are, in a sense,
what a software object guarantees, meaning how the input, output, or how any aspect of the object is guaranteed.
In our interviews, 14 developers gave examples of indirect conflicts they had experienced
which stemmed from not understanding the far reaching ramifications of a change being made to an object contract
towards the rest of the project. Of those 14, 3 dealt
with the changing of legacy code, with one developer saying ``legacy code does not change because developers
are afraid of the long range issues that may arise from that change''. Another developer, in regards to changing
object contracts stated ``there are no changes in the input or changes in the output, but the behavior is different''.
Developers also noted that the conflicts that do occur tend to be quite unique from each other and do not necessarily
follow common patterns.

In regards to object contract changes, 9 developers currently working with large scale database applications listed database
schemas as a large source of indirect conflicts while 5 developers that work on either software
library projects or are in test said that methods or functions were the root of their indirect conflict issues.
7 developers mentioned that indirect conflicts occur when a major update to an external project,
library, or service occurs with one developer noting that ``their build never breaks, but it breaks ours''. Some
other notable indirect conflict artifacts were user interfaces in web development and full components in component
base game architecture.

11 developers explained that indirect conflicts occur ``all the time'' in
their development life cycle with a minimum occurrence of once a week, with more serious issues tending
to occur once a month. The questionnaire showed that, 64\% of indirect conflicts occur bi-weekly
or more frequent,
with 25\% saying that weekly occurrences are most common (seen in Table~\ref{tab:often}). 5 questionnaire participants
also stated that the stage of the development cycle can greatly cause the frequency of indirect conflicts to
differ.

12 developers reported that
when a project is in the early stages of development, indirect conflicts tend to occur far more frequently
than once a stable point is reached. For example, ``At a stable point we decided we are not going to change
[this feature] anymore. We will only add new code instead of changing it.'' and ``the beginning of a project
changes a lot, especially with agile''. From the survey respondents we also learned that ``indirect conflicts after a release
depend on how well the project was built at first'', ``[indirect conflicts] tend to slow down a bit after a
major release, unless the next release is a major rework.'', and ``[indirect conflicts have] spikes during
large revamps or the implementation of cross-cutting new features.'' which corroborate our interview results.
Questionnaire data also showed that
indirect conflicts are more likely to occur before the first major release rather than after at the daily
and weekly occurrence intervals as seen in Table~\ref{tab:often}.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{7cm} | c | c | c | c | c | c | c |}
\hline
Occurrences & Daily & Weekly & Bi-Weekly & Monthly & Bi-Monthly & Yearly & Unknown \\
\hline
\hline
In general & 18\% & 25\% & 21\% & 16\% & 3\% & 5\% & 11\% \\ \hline
Early stages of a development & 32\% & 18\% & 4\% & 5\% & 0\% & 5\% & 36\% \\ \hline
Before the first release & 13\% & 29\% & 6\% & 8\% & 1\% & 3\% & 40\% \\ \hline
After the first release & 6\% & 18\% & 8\% & 18\% & 1\% & 5\% & 44\% \\ \hline
Late stages of development & 6\% & 5\% & 5\% & 18\% & 8\% & 12\% & 46\% \\ \hline
\end{tabular}
\end{center}
\caption{Results of questionnaire as to how often indirect conflicts occur, in terms of percentage
of questionnaire participants.\label{tab:often}}
\end{table*}

In terms of organizational structure, questionnaire data shows that as a project becomes larger and more
developers are added, even to the point that multiple teams are formed, indirect conflicts become more likely to
occur. However, as shown in Table~\ref{tab:env}, indirect conflicts still occur in small teams, with  43\% of respondents indicating
 they are likely to occur in single developer projects.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{7cm} | c | c | c | c | c |}
\hline
Environment of conflicts & Strongly Disagree & Disagree & Neutral & Agree & Strongly Agree \\
\hline
\hline
Developing alone (conflicts in own code) & 18\% & 20\% & 19\% & 24\% & 19\% \\ \hline
Developing in a team between 2 - 5 developers (inter-developers conflicts) & 3\% & 8\% & 22\% & 49\% & 18\% \\ \hline
Developing in a multi team environment (inter-team conflicts) & 1\% & 11\% & 14\% & 39\% & 35\% \\ \hline
\end{tabular}
\end{center}
\caption{Questionnaire results about development environments in which indirect conflicts are likely to occur, in terms of percentage
of questionnaire participants.\label{tab:env}}
\end{table*}

\phantomsection
\addcontentsline{toc}{subsection}{RQ2}
\subsection*{\normalfont{RQ2.} \it{What mitigation techniques are used by developers in regards to indirect conflicts?}}

\bf{Three preventative techniques are used to mitigate indirect conflicts: design by contract, add and deprecate, and personal
experience. For catching indirect conflicts, developers use: testing (unit and integration) with proper use case coverage,
scheduled build processes, and static analysis tools built into IDEs. For resolving indirect conflicts, developers use: historical
error and information logs, compiler and debugger tools, as well as static analysis tools built into IDEs.}

\normalfont{}

Our analysis indicates indicates three major preventative processes used for indirect conflicts.
First, design by contract is heavily used by interviewed developers as a means to avoid indirect conflicts or to understand
when they are likely to occur. The use of design by contract here means that developers tend not to change an object's
contract when possible, and that an object's contract is used as a type of documentation towards awareness of the
software object. One developer stated that ``design by contract was invented to solve this problem and it does it
quite well'', while another noted that software object contracts do solve the problem in theory, but that doesn't
mean that problems don't still occur in practice.
Second, 21\% of developers mentioned that the add and deprecate
model is used to prevent indirect conflicts once
the project, feature, or component has reached a stable or mature point.
Add and deprecate meaning instead of editing code, the developer simply clones old code (if needed), and edits the clone
while slowly phasing out the out the old code in subsequent releases or ad needed. This allows other software to
use the older versions of software objects which remain unchanged, thus avoiding indirect conflicts.
Lastly, pure developer experience was mentioned with 7 developers mentioning that when planning code changes,
either a very experienced member of the project was involved in the planning and has duties to foresee any
indirect conflicts that may arise, or that developers must use their personal knowledge to predict where indirect
conflicts will occur while implementing.

\begin{table*}[tb!]
\begin{center}
\begin{tabular}{| p{7cm} | c | c | c | c | c |}
\hline
Code change type &Never  & Occasionally & Most Times & Always & I Don't Care \\
\hline
\hline
Method signature change & 5\% & 8\% & 12\% & 68\% & 7\% \\ \hline
Pre-condition change & 5\% & 27\% & 37\% & 23\% & 7\% \\ \hline
Main algorithm change & 11\% & 45\% & 19\% & 15\% & 11\% \\ \hline
User interface change & 12\% & 32\% & 20\% & 27\% & 9\% \\ \hline
Run time change & 13\% & 29\% & 25\% & 20\% & 12\% \\ \hline
Post-condition change & 7\% & 28\% & 32\% & 23\% & 11\% \\ \hline
\end{tabular}
\end{center}
\caption{Questionnaire results about source code changes that developers deem notification worthy, in terms of percentage
of questionnaire participants.\label{tab:pre}}
\end{table*}

Of the the 37 questionnaire participants who could give positive identification of preventative processes for indirect conflicts,
27\% said that individual knowledge of the code base and their impact of code change was used, 59\% mentioned some form of design
by contract or the testing of a methods contract, and 14\% said that add and
deprecate was used in their projects to avoid indirect conflicts. This is a confirmation of what was found in the interviews.

In regards to catching indirect conflicts, 13 developers mentioned forms of testing (unit, and integration)
as the major component of catching indirect conflict issues, subscribing to the idea of ``run the regression and integration
tests and just see if anything breaks''. The words ``use case coverage'' were constantly being used by developers
when expressing how proper unit and integration tests should be written. Developers expressed that with proper use case coverage, most if
not all indirect conflicts should be caught. In corroboration,
questionnaire participants corroborated these results with 49\% mentioning forms of testing as the major tool used to
catch indirect conflicts, 33\% said build processes, while 31\% used work their IDE or IDE plug-ins to catch indirect conflicts.
Questionnaire participants also mentioned review process and personal expertise as factors of catching indirect conflicts.

Once an indirect conflict has occurred and developers need to resolve it, 14 developers said
they checked historical logs to help narrow down where the problem could originate from. Most developers had the mindset of
``Look at the change log and see what could possibly be breaking the feature.''. The log most commonly referred to was the source
code change log to see what code has been changed, followed by build failure or test failure logs to examine errors messages and get time
frames of when events occurred. Questionnaire data show that 23\% of developers use native IDE tools and 21\% said they use
features of the language's compiler and debugger in order to solve indirect conflicts. Interestingly, only 13\% of developers
mentioned a form of communication with other developers in aid to resolving these conflicts and only 4\% mentioned the reading
of formal documentation.

%Very important
Through the processes and tools of prevention, detection, and resolution of indirect conflicts, it is important to note that
most developers ascribe to the idea of ``I work until something breaks'', or taking a curative rather than preventative
approach. This means that while developers do have processes and tools for prevention, they would rather spend their time
at the detection and resolutions stages. One developer noted that preventative processes are ``too much of a burden''
while a project manager said ``[with preventative process] you will spend all you time reviewing instead of implementing''.

\phantomsection
\addcontentsline{toc}{subsection}{RQ3}
\subsection*{\normalfont{RQ3.} \it{What do developers want from future indirect conflict tools?}}

\bf{Because developers believe ``good enough'' solutions exist for preventing and detecting indirect conflicts, we identified a strong preference for improvements to debugging processes. Developers prefer automated debugging tools
and better source code analysis tool for impact management and supporting cross team and cross sub-project projects.}

\normalfont{}

When asked about preventative tools, developers had major concerns that the amount of false positives
provided by the tool which may render the tool useless. Developers said ``this would
be a real challenge with the number of dependencies'', ``it depends on how
good the results are in regards to false positives'', and ``I only want to know if it will break me'', meaning that
developers seem to care mostly about negative impacts of code changes as opposed to all impacts in order to reduce
false positives and to keep preventative measures focused on real resulting issues as opposed to preventing potential
issues. Overall, developers had little interest in preventative tools or processes.

In terms of catching indirect conflicts, developers suggested that proper software development processes
are already in place to catch potential issues such as testing, code review,
individual knowledge, or static language analysis tools. Further, developers said that having strict change type monitoring
for indirect conflicts does not work due to the inherent complexities of indirect conflicts.
This can be seen corroborated in that questionnaire participants had very few cases in which they always wanted to be
alerted to a change type as seen in Table~\ref{tab:pre}. Only method signature changes caused 68\% of questionnaire participants
to want to be always notified as they have a high chance to break the code. This is opposed to other change types listed in
Table~\ref{tab:pre} which have 27\% or lower of developers always wanting to be notified. This again showcases the
complexity of indirect conflicts through change types which may or may not negatively affect a project.

When asked about curative tools, developers could only suggest that resolution times be decreased by different means.
Interview and questionnaire participants suggested the following improvements to curative tools:

\begin{itemize}
	\item Aid in debugging by finding which recent code changes are breaking a particular area of code or a test.
	\item Automatically write new tests to compensate for changes.
	\item IDE plug-ins to show how current changes will affect other components of the current project.
	\item Analysis of library releases to show how upgrading to a new release will affect your project.
	\item Built in language features to either the source code architecture (i.e. Eiffel or
				Java Modeling Language~\footnote{http://www.eecs.ucf.edu/~leavens/JML//index.shtml}) or the compile
				time tools to display warning messages towards potential issues.
	\item A code review time tool which allows deeper analysis of a new patch to the project allowing the reviewer to see potential
				indirect conflicts before actually merging the code in.
	\item A tool which is non-obtrusive and integrates into their preexisting development styles without them having to take extra steps.
\end{itemize}

\section{Discussion}
\label{sec:disc}

We now discuss these results in relation to the outstanding issues of tool-based solutions regarding indirect conflicts and which are information overload,
false positives, and scalability.
As per our findings, three main discussion points emerge: (1) What developers really want and don't want from indirect
conflict awareness tools, (2) Whether indirect conflicts are better handled by extensively trying to prevent them or by simply
fixing them as they occur with less prevention, and (3) A gap in analytical tools for software dependencies found in current
industry settings.

\subsection{Unwanted Awareness}

\textit{Developers tend to only care about the awareness of other's activities, if it causes
a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
awareness about most, if not all actions, and it also suggests why developers see a high amount of what they believe to be
false positives, as the changes being reported are not causing them to take action.}

According to our study participants, indirect conflicts represent a serious problem that occur frequently,
sporadically, and differ greatly from case to case. These conditions pose large issues for the creation of generalizable theories
or tools in regards to indirect conflicts. These underlying complexities are the probable cause of disinterest of software developers
to proposed or implemented tools as discussed in Section~\ref{sec:related}. This inability to generalize is what we believe to be the
leading cause of information overload and false positives  awareness systems, causing developers to eventually ignore
information being presented to them, rendering the system useless.

\textbf{These false positives are caused by a difference in what
developers consider to be false positives versus what awareness literature considers they are.} This disjoint, we argue, is caused by developers only considering events which cause some action to be taken on their part, to be true positives,
where as current awareness understanding would state any event which is related to an individual's
work~\cite{Herbsleb06collaborationin}~\cite{Cataldo:2008:SCF}, to be true positives.
``You need a good understanding of what the code change is or else you will have a lot of false positives'' said one developer,
showing that not all changes around an object should be reported for awareness.

Developers have been found to have a great understanding of what they need to know about and more importantly what they don't in
their project awareness. To be able to fully understand a developer's awareness intuition, we can see from the results of this paper that
developers only want awareness of an event if the event forces the developer to take some action.

\textbf{Developers don't care about what they don't need to account for}. ``We would want a high probability of the [reported] change being an issue''.
This sense of unwanted or limited awareness is crucial to understand why generalized
awareness techniques of difficult to generalize problems, such as indirect conflicts through generalized changes to
classes~\cite{Sarma:2007:TSA}, or methods~\cite{Trainer:2005:BGT,Servant:2010:CPI}, often encounter difficulties of false
positives and information overload.
Developers simply do not want to know about events which effect them, but require no action on their part.

This unwanted awareness, or limited awareness, seems counter intuitive to current awareness understanding which would state that
being aware of all events in ones surrounding leads to higher productivity or other quality aspects. In theory this is correct,
but as it was found in practice, this is an incorrect assumption.
In regards to this full awareness, one developer said ``There is no room for this in [our company], as tools are already in place for
analysis of change[s] and code review takes care of the rest''. Since software developers have limits on their time, awareness of
all events surrounding a developer's work or project is not possible.
In terms of developer communication, Hattori et. al.~\cite{Hattori:2012:ICG} found through qualitative user studies, that developers
tend to use the bare amount of communication
towards other developers in order to solve direct conflicts, and take the same approach of only communicating once a conflict has arose.
\textbf{Developers prefer to spend their limited time dealing with
the awareness of events which cause them to take some action (changing code, communication, etc) rather than simply being aware
of events occurring around their work which pose no direct threat to the consumption of their time.}

Of course, whether or not this unwanted awareness is the correct path for developers to take is an open question to consider. When
developers encounter a problem which could have been solved by having greater awareness of events which did not directly affect
them initially, we must consider the positive and negative influences of adding this, for now, unwanted awareness. A positive
influence of total, or near total
awareness of events at the developer level, would be the full understanding of a developer's work and environment which comes
with higher quality or understanding of the product.
A negative influence would be that valuable developer time is spent understanding events which may not directly apply to themselves as opposed to
producing more output of their own work. This balance between awareness and productivity is found to be a fine line in practice,
however, when given the choice, developers tend to lean towards less awareness in order to be more productive in their eyes.

\subsection{Prevention versus Cure}

\textit{Developers would rather spend their time on curative measures (fixing problems after they
arise) than preventative measures (preventing problems from happening). Developers see the task of fixing real problems
to be less of a burden than preventing potential problems because of available tools and their perceived notions of time management.
Focusing on cure is a direct response to the issue of scalability found in indirect conflict tools. Trying to prevent all conflicts
is too large a problem to handle, while curing existing conflicts is manageable by a developer.}

Our findings suggest that developers possess both tools and practices for the prevention,
detection, and resolution of indirect conflicts.

\textbf{Through unwanted awareness, and the use of developer time,
				developers tend to prefer working on real issues that have already occurred as opposed to preventing issues of the future which
				may never arise.} Related to the popular adage ``I work until something breaks'' taken by most developers, this mindset
is a clear example of developers taking a curative approach to software development opposed to a preventative approach.
Prevention here refers to taking precautionary steps to stop potential or some probability of issues, or indirect conflicts, form occurring in the first
place, while cure refers to fixing non probable (have happened) issues as they arise which includes not attempting, or putting little attempt, into preventing
them in the first place.

Two out of the three identified prevention methods taken by developers are simple blanket risk mitigation strategies
accomplished essentially by
not changing code (design by contract, and add and deprecate), while the third is simply developer experience and knowledge. Clearly,
developers are spending little to no time in prevention (or what the perceive as prevention). Developers do however spend a
large amount of time in detection and cure through the
writing of tests and the debugging of issues (although writing tests can be viewed as prevention while their utility is only see at detection).
In fact, most improvements mentioned by developers in regard to dealing with indirect
conflicts occurred at either the detection or cure levels. ``You're reaction time is much more crucial''
said one developer in that resolution tools are believed to be or larger importance as once issues have occurred, it doesn't
matter what prevention was taken, the issue must simply be resolves as quick as possible now.

The lack of prevention process and tools being used is believed to be due to 2 factors as discovered in the interview process.
The first being the identification of dependencies.
Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task
which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts.
Hattori et. al.~\cite{Hattori:2012:ICG} show that with direct conflicts, the
sooner preemptive information if available to developers the more they will communicate and either avoid or easily solve their code merges,
which is the situation many indirect conflict tools have strived for, however the 2 factors of dependencies and knowledge of failure
are what ultimately have led to the amount of false positives and information overload seen in previous tools. \textbf{The
abundance of detection and curative process and tools shows once again developers willingness to debug
real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. The question ultimately
presented in this area is if curative measures are more productive than preventative measures.}

Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
to measure. The pros and cons of prevention versus cure we have identified, are similar to those of unwanted awareness,
in that they result in a trade off
of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
not be thinking of how to move developers into a more preventative mindset. If the inverse is true, should we not be putting more time into automatic debugging, such as
Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.

A last interesting observation about current curative measures being taken in industry, is that developers view testing of software
as curative, when one could easily make the argument that it is preventative in that most tests are written to pass originally and
are kept in place to ensure future changes do not cause issues. This mindset may come from the notion that writing tests is originally
thought of as part of normal code writing, meaning that developers see the extra task of test coverage as part of feature implementation.
This may suggest that if, given the right tools, developers may no longer view preventative
measures as a ``burden'' and may be more inclined to take a more balanced preventative approach rather than mostly curative, as
``prevention is the goal'' was commonly said by developers.

This prevention versus cure discussion resides purely on the developer level and should be noted that it may not apply to system
designers, architects, or managerial stakeholders.
From the project manager's interviewed it was shown that they are heavily favored towards planning and
prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
approaches to their developers.

%mature point, static analysis
\subsection{Gap in Software Evolution Analysis}

\textit{Due to a lack of productive software analysis tools, caused by project infrastructures
and unfriendly analysis languages, contextual indications of indirect conflicts are often missed. This lack of analysis also causes
some software project configurations and software languages to be quite prone to indirect conflicts. This is yet another issue related
to scalability of existing indirect conflict tools.}

Indirect conflicts are more likely to occur,
from developer opinion, before a mature point is reached in the project's evolution. ``Once you get the API stable, people are
better at communicating changes in regards to dependency concerns.''(This mature point may stem from a major release,
end of an iteration, a new feature being released or any point considered stable and reliable) However, this context of a
mature point, as well as any other
potential contextual attributes, are rarely identified or accounted for in regards to indirect conflicts. Developers have said
that different change types may occur at different rates throughout a project's life time and that this
may drastically effect the outcomes of indirect conflict tools or processes.

\textbf{A deeper understanding of what context indirect conflicts occur in seems to be more of a success factor to indirect conflict
research than may have been previously thought.} Static analysis may be useful in regards to indirect conflict context in that we
could identify trends surrounding the mature points of previous projects in order to give a better understanding of what it
means for a project to be pre or post mature point, which could affect the outcome of indirect conflict understanding.
Fluri et al.~\cite{Fluri:2007:CDT} uses static analysis and abstract syntax trees to be able to determine what types of source code
changes occur at a given point in time which could be used for determining such a context of mature points.

This added level of context,
would add to the prevention versus cure debate as previously discussed. Prevention may be a better
choice once a project has reached a mature point as the code base becomes more stable and source code changes become more dangerous
to the quality of the project. Curative measures may be a better choice before a mature point as code churn can be higher which
causes more bugs than can be prevented. It may be imperative to discover more on project stability
and the mature point in order to fully understand the nature of indirect conflicts and their context.

While this added sense of understanding is important to indirect conflicts, an understanding of a project's evolution,
a mature point being reached as an example,
may pose as a more useful tool for project management rather than developers. This more abstract tendency of indirect
conflict occurrences may add even more power to project management for evaluation of progress, code stability, and code reviews.

Aside from context, dependency identification and tracking is a key missing component of
indirect conflict analysis due to the weaknesses of static analysis. The gap in this identification comes from software and
organizational structures of software teams. \textbf{Between increased modularization (multiple sub projects or repositories),
cross language dependencies, and languages which do not lend themselves to static analysis~\cite{Jensen:2011:MHD}, static analysis tools have
become quite limited in identifying and tracking dependencies where they were once strong.}
In a study regarding static analysis tools, Johnson et. al~\cite{Johnson:2013:WDS} found similar conclusions towards the
ill fit of current static analysis to current development practices..

Relational database schemas are one of the highest sources of indirect
conflicts found in their projects, ``it breaks stuff all over the place'',
yet we know from Maule et al.~\cite{Maule:2008:IAD} that relational database schemas have
been scarcely researched in terms of indirect conflicts. This falls in with our previous understanding of cross language
dependency tracking in that database schemas are independent of languages which may be on the receiving end of their output.

These increases of complexity in software products has left quite a gap in dependency identification and tracking which
has lead to some of the deficiencies of indirect conflict research.

\subsection{Implications for Future Research Work}
\label{sec:implr}

\textbf{Prevention versus Cure} \textit{There is a need for continued study of the open question
``Which of prevention or cure is more effective for software development and indirect conflicts?''. We believe that this simple question will undoubtedly
produce extremely complex answers. Developers are currently favoring cure, but we should be mindful that what developers want
may not always be the best choice.}

The fact that software developers focus their current efforts on curative measures, for
indirect conflicts, suggests
that while it may not be the most effective, it may be the easiest road for developers to take. This may be the path of
least resistance, but it may not be optimal. Software engineering as a whole should strive to answer this question or
provide more insight into possibilities, as its answer may determine many future actions taken by the research community.

Recommendations towards the study of prevention versus cure involve the examination of formal processes and tools used
by industry professionals with measurements of efficiency and effectiveness, similar to the work of
Tiwana~\cite{Tiwana:2008:ICD} in coordination tools. With these studies, we may find insight into the correct balance
of prevention versus cure, thus being able to increase developer productivity as well as identify more gaps in theory
versus practice which may lead to improved tools or the abandoning of existing ones as was shown with UML~\cite{Petre:2013:UP}.

\textbf{Theories of Awareness to Model Real World} \textit{There is a need to further characterize the
mismatch between awareness approaches and tools as developed in the research community and awareness needs as perceived
by industry. Theories of awareness should consider modeling real development
environments where interruptions, time management, and full development life cycles are often large factors.}

As was identified in the previous discussion, while current understanding suggests that awareness of all events related to ones work
produces a more coherent understanding of a person's environment, developers find this to be overly time consuming
to the point where they only want to be aware of events which require action on their part. This difference, of what
should be and what is, should be further understood to combat future potential failures in
tool development or theories attempted to be used in practice.

\subsection{Implications for Future Tool Work}
\label{sec:implt}

\textbf{Give Developers What They Want} \textit{Developers have their own notions of how ``good''
something is for their productivity. We should pay attention to what they want, in order to shape our
models and better address their needs.}

The direct implication this research has on tool development is that of current adoption among developers. As was stated,
developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being
said, focus at these two stages for tool developer will lead to larger adoption among developers.
It should be noted here that detection must come with an almost zero rate of false positives as
current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.

\textbf{Stronger Source Code Analysis} \textit{With languages like JavaScript becoming popular, having projects with
multiple languages, and more modular structures, we should focus on better techniques for static analysis based
on what industry standards.}

The more indirect implication of this research on tool creation is that of improving existing tools. While not all existing
tools are used for indirect conflicts alone (automatic debugging~\cite{Zeller:2005:WPF}, unit tests, etc), most of these tools
have a need for rapid expansion, according to developers, for dealing with indirect conflicts. The ability to have unit tests
automatically written for a given software object's contract, the ability to find a change  in an external project
which has broken a developer's own project, or any automation of the existing detection and resolution stages of indirect
conflicts are what developers currently seek. But of course, most of these implications rely on the improvement of
static analysis tools.

These tool implications themselves imply the need of further development of static analysis tools. Static
analysis lays at the heart of most if not all stages of indirect conflict research. We must be able to track and manage
software dependencies across the new landscape of software development that is multiple projects, repositories, and cross
language support. These improvements will allow the further development of both current and future indirect conflict
tools.

\section{Evaluation and Threats to Validity}
\label{sec:eval}

From the re-interviewed participants, we found extremely positive feedback regarding both our results and major discussion
points. Participants often had new stories and experiences to share once they had heard the results of this paper which
confirmed the findings and often were quite shocked to hear the results as they did not usually think about their actions
as such but then realized the results held true to their daily actions for better or worse.

As per grounded theory research, Corbin and Strauss list ten criteria to evaluate quality and credibility~\cite{Corbin:1998:SP}.
We have chosen three of these criteria and explain how we fulfill them. The three evaluation criteria selected are more relevant
to a qualitative study than the usual threats to validity associated with qualitative work. These selected criteria can also
be viewed as the threats to validity of this paper.

{\bfseries Fit.} ``Do the findings fit/resonate with the professional for whom the research was intended and the participants?'' This
criterion is used to verify the correctness of our finding and to ensure they resonate and fit with participant opinion. It is also
required that the results are generalizable to all participants but not so much as to dilute meaning. To ensure fit, during interviews
after participants gave their own opinions on a topic, we presented them with previous participant opinions and asked them to comment
on and potentially agree with what the majority has been on the topic. Often the developers' own opinions already matched those of
the majority before them and did not necessarily have to directly verify it themselves.

As added insurance, we conducted 5 confirmatory interviews after we analyzed our data and drafted themes in our findings. We describe this
procedure in Section~\ref{sec:meth}.

To ensure the correctness of the results, we also linked all findings in Section~\ref{sec:results} to either a majority of agreeing
responses on a topic or to a large amount of direct quotes presented by participants.

{\bfseries Applicability or Usefulness.} ``Do the findings offer new insights? Can they be used to develop policy or change practice?''
Although our results may not be entirely novel or even surprising, our findings do (1) suggest
a disjoint between the current research understanding of developers' awareness needs and the these awareness needs related to indirect conflict tools in practice, and (2) provide some
insight into the debate of prevention versus cure in software development. Given that many indirect conflict tools still suffer from these same common
issues, we believe that these findings will help researchers focus on what developers want and need moving into the future more than has
been possible in the past. These finding set a course of action for where effort should be spent in academia to better benefit industry.

%10 of the 78 questionnaire participants sent direct responses to us asking for any results of the research to be sent directly to
%them in order to improve their indirect conflict work flows. 7 of the 19 participants interviewed expressed interest concerning any possible
%tools or plans for tools inspired by this research as well. The combination of academia relatability and direct industry interest
%in our results help us fulfill this criterion.

{\bfseries Variation.} ``Has variation been built into the findings?'' Variation shows that an event is complex and that any findings
made accurately demonstrate that complexity. Since interviewed participants came from such a diverse set of organizations,
software language knowledge, and experience the variation naturally reflected the complexity. Often in interviews and questionnaires, participants
expressed unique situations that did not fully meet our generalized findings or on going theories. In these cases, we worked in the specific
cases which were presented as boundary cases and can be seen in Section~\ref{sec:results} as some unique findings or highly
specialized cases. These cases add to the variation
to show how the complexity of the situation also resides in a significant number of unique boundary situations as well as the complexity
in the generalized theories and findings.

\section{Conclusions}
\label{sec:conc}

Indirect conflicts are a significant issue with real world development, however, many of our research proposed techniques and tools have been unsuccessful in attracting major support from developers. Based on our qualitative study involving
19 interviewed developers from 12 organizations as well as 78 questionnaires respondents, we sought to characterize indirect conflicts,
current developer work flows surrounding indirect conflicts, and frame directions for future tool development that support developers
in their tackling of indirect conflicts.

Our findings  indicate that indirect conflicts occur frequently and are likely caused by software contract changes;
while design by contract, add and deprecate, and personal experience help prevent indirect conflicts,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, and developers want indirect conflict tools to focus on automatic debugging and better source code analysis.

Our study of practitioners intends to fill a gap in the software engineering research with respect to the outstanding problems in current tool-based solutions to indirect conflicts, namely the issues of information overload, false positives and scalability. Overall, we described a disjoint between the current research understanding of developers' awareness needs and these awareness needs related to indirect conflict tools in practice. More specifically, our findings suggest that developers do not want awareness mechanisms which provide non actionable results; developers
would rather focus on curing existing problems rather than preventing potential issues;
and that there exists a gap in software evolution analytical tools between what is available and what practitioners need. We hope that these insights will be found valuable in future research endeavours and development of tools to assist developers in tackling indirect conflicts.

\bibliographystyle{IEEEtran}
%\balance
\bibliography{paper}

% End of the paper
\end{document}