### 2014-03-17 msc-thesis/chapters/1/chapter-intro.tex
MSC final melanie edits.

\startfirstchapter{Introduction}
\label{chapter:introduction}

\section{Introduction}

% Introduce awareness
As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness
mitigates some downsides of parallel development which include the cost of conflict prevention and resolution. However,
empirical evidence shows that these mitigated losses continue to appear quite frequently and can prove to be a significant
and time-consuming chore for developers~\cite{Perry:2001:PCL}.

% Introduce dependencies
Large software projects are created using highly modular and reusable code. This creates technical dependencies between
methods or functions that can be used in a wide variety of locations throughout the project. This causes changes to any 
given software object to have a rippling effect across the rest of the project~\cite{Acharya:2011:PCI}. The larger these 
effects are, the more likely they are to cause a software failure inside the system during the project's life 
span~\cite{Zimmermann:2008:PDU}. These observations of technical dependencies open the door to types of analysis on the 
developer networks they infer and preventing software failures by improving coordination amongst dependent developers.

% What has been done with technical dependencies
Technical dependencies in a project can be used to predict success or failure in builds or code 
changes~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU}. However, most research in this area is based on identifying central 
modules inside a large code base which are likely to cause software failures or detecting frequently changed code that 
can be associated with previous failures~\cite{Kim:2006:AIB}. This module-based method also results in predictions at 
the file or binary level of software development as opposed to a code change level and often lack the ability to provide 
recommendations for improved coordination other than test focus.

% Using technical dependencies for awareness tools
With the power of technical dependencies in tracking unintended consequences from source code changes, several tools 
have been created to attempt to solve task awareness related issues with some 
success~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC}. However, these tools have been designed 
to solve task awareness related issues at the direct conflict level.

%Describe difference between direct and indirect
Two types of conflicts have attracted the attention of researchers, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact. Tools have been created and studied for direct conflicts
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. Indirect conflicts are caused by source code changes that negatively impact another
location in the software system such as when libraries are upgraded or when a method is changed
and invoking methods are influenced negatively. Indirect conflict tools however, have
not shared the same success as direct conflict 
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. However,
previous interviews and surveys conducted with software developers have shown a pattern that developers 
of a software project view indirect conflict awareness  as a high priority issue in their 
development~\cite{Damian:2007:GSE, Halverson:2006:DTV, Begel:2010:CDE, Schroter:2012:TTF}, meaning that future research
is required to address this developer concern.

% Limitations of IC tools for indirect conflicts compared to Impact
Indirect conflicts arising in source code are inherently difficult to resolve as most of the time, source code analysis must 
be performed in order to find relationships between technical objects which are harmed by changes. While some awareness 
tools have been created with these indirect conflicts primarily in mind~\cite{Begel:2010:CDE, Trainer:2005:BGT}, most have only 
created an exploratory environment which is used by developers to solve conflicts which may arise. These tools were not designed 
to detect indirect conflicts that arise and alert developers to their presence  inside the software system. 
Sarma et al.~\cite{Sarma:2007:TSA} has started to work directly on solving indirect conflicts, however, these products are not 
designed to handle internal structures of technical objects.

% Issues with IC tools of information overload etc.
While indirect conflict tools have shown potential from studies of developers, some of the same problems continue
to arise throughout most, if not all tools. The most prevalent issue is that of false positives and information
overload, tools eventually being
ignored~\cite{Sarma:2007:TSA, Servant:2010:CPI}. A second primary issue is that of dependency identification and
tracking. Many different dependencies have been proposed and used in indirect conflict tools such as method 
invocation~\cite{Trainer:2005:BGT}, and class signatures~\cite{Sarma:2007:TSA} with varying success, but the 
identification of failure inducing changes, other than those which are already identifiable by other means such
as compilers, and unit tests, to these dependencies still remains an issue. Dependency tracking issues are
also compounded by the scale of many software development projects leading to further information overload.
Lastly, social factors such as Cataldo et al's.~\cite{Cataldo:2006:ICR} notion of socio-technical
congruence have been leveraged in indirect conflict tools~\cite{Kwan:2011:ESC, Begel:2010:CDE, Borici:2012:CHA}.
However, issues again of information overload, dependencies (in developer organizational structure), and scalability 
come up.

% What is being addressed in this thesis
Clearly, indirect conflicts and its subsequent research areas have a large breadth of limitations, some of which
will be explored in this dissertation. The research goal of this dissertation is to explore the limitations which exist
with supporting indirect conflicts through awareness techniques as well as to explore possible solutions for industry practice 
in the area of indirect conflicts. To accomplish this goal, I have researched the following sub topics of indirect
conflicts: technical dependencies, how developers are involved in said dependencies,
socio-technical congruence as a mitigation strategy to indirect conflicts, what the root causes of indirect conflicts are,
what compounding factors exist for indirect conflicts, what current industry mitigation strategies of indirect conflicts
are being used, what future steps should be taken by researchers to better industry regarding indirect conflicts, and finally,
how software evolution analysis can be used to better tools for indirect conflicts. I have addressed these issues by
conducting four studies.

\section{Research Methodology}

In order to address the research goal as laid out in the previous section, I have conducted 4 studies which will
now be briefly outlined. Each study builds off the previous one and has research questions informed from the
findings of the previous study.

% Study 1 introduction
Study 1 focuses on the power of technical dependencies in software projects. The question I
investigated were : ``\textit{Is it possible to identify pairs of developers whose technical dependencies in 
code changes statistically relate to bugs?}''. This study explains the approach used to locate these pairs of developers 
in developer networks.The process utilizes code changes and the call hierarchies effected to find patterns of developer 
relationships in successful and failed code changes. These developer pairs can be seen as indirect conflicts occurring
as one developer's code change has negatively affects another developer's work. As it will be seen, I found 27 statistically significant failure 
inducing developer pairs. These developer relationships can be used to promote the idea of leveraging socio-technical 
congruence, a measure of coordination compared to technical dependencies amongst stakeholders, to provide coordination recommendations.
This notion of socio-technical congruence is my initial proposed solution to indirect conflicts. By identifying these failure
inducing pair of developers over indirect conflicts, I hoped that recommended communication would be the correct fix. The 
results of Study 1 directly influence Study 2.

% Study 2 introduction
Study 2 attempts to take the failure inducing pairs of developers from Study 1 and create an awareness tool while
answering: ``\textit{Can indirect conflicts be supported through an awareness mechanism which leverages pairs of
developers whose changing technical dependencies statistically relate to bugs?}''.
In this study, I report on my research into supporting indirect conflicts and present the design, implementation, and 
evaluation of the tool \textit{Impact}, a web based tool that aims at detecting indirect conflicts among developers and 
notifying the appropriate members involved in these conflicts. By leveraging technical relationships inherent of software 
projects with method call graphs~\cite{Lakhotia:1993:CCM} as well as detecting changes to these technical relationships 
through software configuration management (SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as 
alert developers involved in such conflicts in task awareness. While this study outlines \textit{Impact's} specific 
implementation, its design is rather generic and can be implemented in similar indirect conflict awareness tools. 
\textit{Impact} represents a first step towards the design and implementation of awareness tools which address indirect 
conflicts in software development. After a brief evaluation of \textit{Impact} with two student software teams, it was
found that \textit{Impact} suffers from information overload and a high false positive rate which turn out to be quite
large problems found in many other indirect conflict
tools~\cite{Sarma:2007:TSA, Holmes:2010:CAR, Trainer:2005:BGT, Servant:2010:CPI, Borici:2012:CHA}. In order to fully understand 
the causes of these indirect conflict tool issues, a third study was conducted.

% Study 3 introduction

In order to fully understand the root causes of information overload, false positives, and
scalability issues in regards to indirect conflicts, Study 3 was an empirical study to determine what events occur to
cause indirect conflicts, when they occur, and if conditions exist to provoke more of these events. 
I also set out to understand what mitigation strategies developers currently use as opposed to those created
by researchers. Through this exploration, I examined what can be accomplished moving forward with indirect conflicts
in both research and industry. This study asked the following 3 research questions: 
\textit{What are the types, factors, and frequencies of indirect conflicts?}
\textit{What mitigation techniques are used by developers in regards to indirect conflicts?}
\textit{What do developers want from future indirect conflict tools?}

I interviewed 19 developers from across 12 commercial and open source projects, followed by a confirmatory survey of 78 
developers, and 5 confirmatory interviews, in order to answer the aforementioned questions. The study findings indicate that: 
indirect conflicts occur frequently and are likely caused by software contract changes and a lack of understanding of those changes,
developers tend to prefer to use detection and resolution processes or tools
over those of prevention, developers do not want awareness mechanisms which provide non actionable results, 
and that there exists a gap in software evolution analytical tools arising from the reliance on static analysis resulting in missed
context of indirect conflicts. As a result of the final finding (the gap in software evolution analytical tools), I conducted
a fourth and final study.

% Study 4 introduction
In order to begin to address the gap in software evolution analytical tools discovered in study 3, I turn my analysis to the notions 
of software change trends, specifically those trends around major releases. Change trends are trends which indicate
a likelihood for a change type to occur around a certain event. Change trends have been used to detect
stability in core architecture~\cite{Wermelinger:2008:AEE} as well as evolving dependencies~\cite{Businge:2010:ESE}.
With the power of major release points in open source projects as a starting point for project stability and the understanding that change trends can
be leveraged to detect stability and the proneness of indirect conflicts (as will be seen in Study 3), this study investigates the question:
``\textit{What trends exist in source code changes surrounding major releases of open source projects as a notion towards a project
stability measure?}''. I perform a case study of 10 open source projects in order to study their source code change trends surrounding major release points
throughout their history. I studied 26 change trends quantitatively and 4 change trends qualitatively, and identified a core group of 9 change trends which occur
prominently at major release points of the projects studied.

% The rest of the thesis 
The remainder of this dissertation is laid out as follows. Chapter two includes Study 1 and Study 2 as the motivational studies which
ultimately led to the larger research studies found in Study 3 and Study 4. Chapter 3 includes Study 3 and Study 4 which ultimately
press upon indirect conflicts in a more in depth fashion than has been previously seen in research. Chapter 4 contains a lengthy discussion
of what has been learned from all four studies of this dissertation as well as implications for further research and tool development
in the field of indirect conflicts. Finally, Chapter 5 concludes this dissertation.